QPLIB: A Library of Quadratic Programming Instances

### 2017-02-01 QPLIB-1.0.tex
update to qplib format and that pesky 1/2

The structure of this paper is the following. In \S~\ref{sec:QPbasic} we review the basic notions about QP. In particular, \S~\ref{subsec:notation} sets out the notation, \S~\ref{sec:classification} proposes a---to the best of our knowledge, new---taxonomy of QP that helps us in discussing the (very) different classes of QPs, \S~\ref{sec:algo} very briefly reviews the solution methods for QP and the solvers we have employed.
Then, \S~\ref{sec:lib} describes the process used to obtain the library and its results.
%; the software tools that we have used, and that are freely released together with the library, are discussed in \S~\ref{subsec:tools}. 
Some conclusions are drawn in \S~\ref{sec:conclusions}, after which Appendix A provides a complete description of all the instances of the library, while Appendix B describes the QPLIB file format.

While no performance issues of solvers for QP/QCP problems are considered in 
this paper, we refer to the comprehensive benchmark site \url{http://plato.asu.edu/bench.html}. Of the benchmarks
on this site three deal exlusively with QP/QCP problems: 
the (1) Large SOCP, (2) MISOCP, and the (3) MIQ(C)P benchmarks. 
The following contain partly results for such problems: 
(4) Parallel Barrier Solvers on Large LP/QP problems, (5) AMPL-NLP, (6) MINLP.
The benchmarks (1,2,4) have only convex instances, while the others include 
nonconvex instances. Global optima are obtained by several of the solvers in
the benchmarks (3,5).
All solvers in the latest addition (6) compute global optima, they are ANTIGONE,
BARON, COUENNE, LINDO, and SCIP, c.f.~\S~\ref{subsec:solver}. This benchmark is based on MINLPLib~2 \cite{Vigerske2014}, a
collection of presently 1388 instances. In order to give a first representative
impression of solver performance, care was taken to reduce the number of instances and 
allow all solvers to finish in reasonable time. More than half of the selected
instances are QP or QCP. A time limit of two hours on a
modern workstation with sufficient memory was used. Computations were done in
GAMS with scripts derived from the SCIP check scripts and available at GitHub.
For details we refer to \url{http://plato.asu.edu/ftp/minlp.html}.

>>>>

The structure of this paper is the following. In \S\ref{sec:QPbasic} we review the basic notion of QP. In particular, \S\ref{subsec:notation} sets out the notation, \S\ref{sec:classification} proposes a---to the best of our knowledge, new---taxonomy of QP that helps us in discussing the (very) different classes of QPs, \S\ref{sec:algo} very briefly reviews the solution methods for QP and the solvers we have employed.
Then, \S\ref{sec:lib} describes the process used to obtain the library and its results.
%; the software tools that we have used, and that are freely released together with the library, are discussed in \S\ref{subsec:tools}.
Some conclusions are drawn in \S\ref{sec:conclusions}, after which Appendix A provides a complete description of all the instances of the library, while Appendix B describes the QPLIB file format.

While no performance issues of solvers for QP/QCP problems are considered in
this paper, we refer to the comprehensive benchmark site \url{http://plato.asu.edu/bench.html}. Of the benchmarks
on this site three deal exlusively with QP/QCP problems:
the (1) Large SOCP, (2) MISOCP, and the (3) MIQ(C)P benchmarks.
The following contain partly results for such problems:
(4) Parallel Barrier Solvers on Large LP/QP problems, (5) AMPL-NLP, (6) MINLP.
The benchmarks (1,2,4) have only convex instances, while the others include
nonconvex instances. Global optima are obtained by several of the solvers in
the benchmarks (3,5).
All solvers in the latest addition (6) compute global optima, they are ANTIGONE,
BARON, COUENNE, LINDO, and SCIP, c.f.~\S\ref{subsec:solver}. This benchmark is based on MINLPLib~2 \cite{Vigerske2014}, a
collection of presently 1388 instances. In order to give a first representative
impression of solver performance, care was taken to reduce the number of instances and
allow all solvers to finish in reasonable time. More than half of the selected
instances are QP or QCP. A time limit of two hours on a
modern workstation with sufficient memory was used. Computations were done in
GAMS with scripts derived from the SCIP check scripts and available at GitHub.
For details we refer to \url{http://plato.asu.edu/ftp/minlp.html}.

### 2017-02-01 QPLIB-2.1.tex
update to qplib format and that pesky 1/2

In mathematical optimization, a Quadratic Program (QP) is an optimization problem in which either the objective function, or some of the constraints, or both, are quadratic functions. More specifically, the problem has the form
%
\begin{align*}
 \min \textrm{or} \max \;\;
 & x^\top Q^0 x + b^0 x + q^0 \\
 \textrm{such that}\;\;
 & c^i_l \leq x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\
 %
 & l_j \leq x_j \leq u_j & j \in \Mcn  \\
 %
 & x_j \in \mathbb{Z} & j \in \Mcz
\end{align*}

%
%\begin{align*}	
% \min \;\;
% & x^\top Q^0 x + b^0 x \\
% %
% & x^\top Q^i x + b^i x \leq c^i & i \in \Mcm \\ 
% %
% & l_j \leq x_j \leq u_j & j \in \Mcn  \\
% %
% & x_j \in \mathbb{Z} & j \in \Mcz
%\end{align*}
%
where:
%
\begin{itemize}
 \item $\Mcn = \{ 1, \ldots, n \}$ is the set of (indices) of variables, and $\Mcm = \{ 1, \ldots, m \}$ is the set of (indices) of constraints;
 %
 \item $x = [x_j]_{j = 1}^n \in \RR^n$ is a finite vector of real variables;
 %
 \item $Q^i$ for $i \in \{ 0 \} \cup \Mcm$ are symmetric $n \times n$ real matrices: because one is always only interested in the value of quadratic functions of the type $x^\top Q^i x$, symmetry can be assumed without loss of generality by just replacing both $Q^i_{hk}$ and $Q^i_{kh}$ with their average $(Q^i_{hk} + Q^i_{kh}) / 2$;
 %
 \item $b^i$, $c_u^i$, $c_l^i$ for $i \in \{ 0 \} \cup \Mcm$, and $q^0$ are, respectively, real $n$-vectors and real constants;
 %
 \item $-\infty \leq l_j \leq u_j \leq \infty$ are the (extended) real upper and lower bounds on each variable $x_j$ for $j \in \Mcn$;
 %
 \item $\Mcm = \Mcq \cup \Mcl$ where $Q^i = 0$ for all $i \in \Mcl$ (i.e., these are the linear constraints, as opposed to the truly quadratic ones);
 %
 \item the variables in $\Mcz \subseteq \Mcm$ are restricted to only attain integer values.
\end{itemize}
%
Due to the presence of integral requirements on the variables and quadratic constraints, this class of problems is often referred to as Mixed-Integer Quadratically Constraint Quadratic Program (MIQCQP). It will be sometimes useful to refer to the (sub)set $\Mcb =  \{ \, i \in \Mcz : l_j = 0 , u_j = 1 \, \} \subseteq \Mcz$ of the binary variables, and to $\Mcr = \Mcn \setminus \Mcz$ as the set of continuous ones. Similarly, it will be sometimes useful to distinguish the (sub)set $\Mcx = \{ \, j : l_j > -\infty \lor u_j < \infty \, \}$ of the box-constrained variables from $\Mcu = \Mcn \setminus \Mcx$ of the unconstrained ones (in the sense that finite bounds are not explicitly provided in the data of the problem, although they may be implied by the other constraints).

>>>>

In mathematical optimization, a Quadratic Program (QP) is an optimization problem in which either the objective function, or some of the constraints, or both, are quadratic functions. More specifically, the problem has the form
%
\begin{align*}
 \min \textrm{or} \max \;\;
 & \half x^\top Q^0 x + b^0 x + q^0 \\
 \textrm{such that}\;\;
 & c^i_l \leq \half x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\
 %
 & l_j \leq x_j \leq u_j & j \in \Mcn  \\
 %
 & x_j \in \mathbb{Z} & j \in \Mcz
\end{align*}

%
%\begin{align*}
% \min \;\;
% & x^\top Q^0 x + b^0 x \\
% %
% & x^\top Q^i x + b^i x \leq c^i & i \in \Mcm \\
% %
% & l_j \leq x_j \leq u_j & j \in \Mcn  \\
% %
% & x_j \in \mathbb{Z} & j \in \Mcz
%\end{align*}
%
where:
%
\begin{itemize}
 \item $\Mcn = \{ 1, \ldots, n \}$ is the set of (indices) of variables, and $\Mcm = \{ 1, \ldots, m \}$ is the set of (indices) of constraints;
 %
 \item $x = [x_j]_{j = 1}^n \in \RR^n$ is a finite vector of real variables;
 %
 \item $Q^i$ for $i \in \{ 0 \} \cup \Mcm$ are symmetric $n \times n$ real (Hessian) matrices: since one is only interested in the value of quadratic forms of the type $x^\top Q^i x$, symmetry can be assumed without loss of generality by just replacing off diagonal pairs $Q^i_{hk}$ and $Q^i_{kh}$ with their average $(Q^i_{hk} + Q^i_{kh}) / 2$;
 %
 \item $b^i$, $c_u^i$, $c_l^i$ for $i \in \{ 0 \} \cup \Mcm$, and $q^0$ are, respectively, real $n$-vectors and real constants;
 %
 \item $-\infty \leq l_j \leq u_j \leq \infty$ are the (extended) real upper and lower bounds on each variable $x_j$ for $j \in \Mcn$;
 %
 \item $\Mcm = \Mcq \cup \Mcl$ where $Q^i = 0$ for all $i \in \Mcl$ (i.e., these are the linear constraints, as opposed to the truly quadratic ones);
 %
 \item the variables in $\Mcz \subseteq \Mcm$ are restricted to only attain integer values.
\end{itemize}
%
Due to the presence of integral requirements on the variables and quadratic constraints, this class of problems is often referred to as Mixed-Integer Quadratically Constraint Quadratic Program (MIQCQP). It will be sometimes useful to refer to the (sub)set $\Mcb =  \{ \, i \in \Mcz : l_j = 0 , u_j = 1 \, \} \subseteq \Mcz$ of the binary variables, and to $\Mcr = \Mcn \setminus \Mcz$ as the set of continuous ones. Similarly, it will be sometimes useful to distinguish the (sub)set $\Mcx = \{ \, j : l_j > -\infty \lor u_j < \infty \, \}$ of the box-constrained variables from $\Mcu = \Mcn \setminus \Mcx$ of the unconstrained ones (in the sense that finite bounds are not explicitly provided in the data of the problem, although they may be implied by the other constraints).

### 2017-02-01 QPLIB-2.2.tex
update to qplib format and that pesky 1/2

Despite the apparent simplicity of the definition given in \S~\ref{subsec:notation}, Quadratic Programming instances can be of several rather different ``types'' in practice, depending on the fine details of the data. In particular, many algorithmic approaches can only be applied to QP when the data of the problem has specific properties. A taxonomy of QP instances should therefore strive to identify the set of properties that an instance should have in order to apply the most relevant computational methods. However, the sheer number of different existing approaches, and the fact that new ones are continuously proposed, makes it hard to provide a taxonomy that is both simple and covers all possible special cases. This is why, in this paper, we propose an approach that aims at finding a good balance between simplicity and coverage of the main families of computational methods.

@@

We will now provide a first general discussion about the different problem classes that our proposed taxonomy defines.
For simplicity, we will assume minimization problems for the remaining of this section.
Some problem classes are actually ``too simple'' to make sense in our context. For instance, D*B problems have only diagonal quadratic (hence separable) objective function and bound constraints; as such, they read
\[
 \textstyle
 \min \; \Big\{ \,
 \sum_{j \in \Mcn} \big( \, Q^0_j x_j^2 + b^0_j x_j \, \big) \;:\;
 l_j \leq x_j \leq u_j \quad j \in \Mcn \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcz \, \Big\}
 \;\; .
\]
Hence, their solution only requires the independent minimization of a convex quadratic univariate function in each single variable $x_j$ over a box constraint and possibly integrality requirements, which can be attained trivially in $O(1)$ (per variable) by closed-form formul{\ae}, projection and rounding arguments. \emph{A fortiori}, the even simpler cases L*B, D*N and L*N (the latter obviously unbounded unless $b^0 = 0$) will not be discussed here. Similarly, CCN are immediately solved by linear algebra techniques, and therefore are of no interest in this context. On the other end of the spectrum, in general QP is a hard problem. Actually, LIQ---linear objective function and quadratic constraints in integer variables with no finite bounds, i.e. 
\[
 \textstyle
 \min \; \Big\{ \, b^0 x \;:\;
 x^\top Q^i x + b^i x \leq c^i \quad i \in \Mcm \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcn \, \Big\}
 \;\; ,
\]
is not only $\mathcal{NP}$-hard, but downright undecidable \cite{jeroslow}. Hence so are the ``harder'' \{C,Q\}IQ.

@@

Another relevant reformulation trick concerns the fact that, as soon as quadratic constraints are allowed, then a linear objective function can be assumed w.l.o.g.. Indeed, any Q** (C*C) problem can always be rewritten as
%
\begin{align*}	
 \min \;\;
 & x^0 \\
 %
 & -\infty \leq x^\top Q^0 x + b^0 x \leq x^0 & \\ 
 %
 & c^i_l \leq x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\ 
 %
 & l_j \leq x_j \leq u_j & j \in \Mcn  \\
 %
 & x_j \in \mathbb{Z} & j \in \Mcz
\end{align*}
%
i.e., a L*Q (L*C). Hence, it is clear that quadratic constraints are, in a well-defined sense, the most general situation (cf.~also the result above about hardness of LIQ).

When a $Q^i$ is positive semidefinite (PSD), i.e., the corresponding constraint/objective function is convex, general quadratic constraints are in fact equivalent to diagonal ones. In fact, since every PSD matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$. Hence, one could think that D** problems need not be distinguished from C** ones; however, this is true only for ``complicated'' constraints, but not for ``simple'' ones, because the above reformulation technique introduces linear constraints. Indeed, while C*L (and, a fortiori, C*\{C,Q\}) can always be brought to D*L (D*\{C,Q\}), using the same technique C*B becomes D*L, which is significantly different from D*B. In practice, a diagonal convex objective function under linear constraints is found in many applications (e.g., \cite{FrFG16,FrGe06a,FrGe09a,FGGP11}), so that D*L still makes sense to distinguish the case where the objective function is ``naturally'' separable from that where separability is artificially introduced.

>>>>

Despite the apparent simplicity of the definition given in \S\ref{subsec:notation}, Quadratic Programming instances can be of several rather different ``types'' in practice, depending on the fine details of the data. In particular, many algorithmic approaches can only be applied to QP when the data of the problem has specific properties. A taxonomy of QP instances should therefore strive to identify the set of properties that an instance should have in order to apply the most relevant computational methods. However, the sheer number of different existing approaches, and the fact that new ones are continuously proposed, makes it hard to provide a taxonomy that is both simple and covers all possible special cases. This is why, in this paper, we propose an approach that aims at finding a good balance between simplicity and coverage of the main families of computational methods.

@@

We will now provide a first general discussion about the different problem classes that our proposed taxonomy defines.
For simplicity, we will assume minimization problems for the remaining of this section.
Some problem classes are actually ``too simple'' to make sense in our context. For instance, D*B problems have only diagonal quadratic (hence separable) objective function and bound constraints; as such, they read
\[
 \textstyle
 \min \; \Big\{ \,
 \sum_{j \in \Mcn} \big( \, \half Q^0_j x_j^2 + b^0_j x_j \, \big) \;:\;
 l_j \leq x_j \leq u_j \quad j \in \Mcn \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcz \, \Big\}
 \;\; .
\]
Hence, their solution only requires the independent minimization of a convex quadratic univariate function in each single variable $x_j$ over a box constraint and possibly integrality requirements, which can be attained trivially in $O(1)$ (per variable) by closed-form formul{\ae}, projection and rounding arguments. \emph{A fortiori}, the even simpler cases L*B, D*N and L*N (the latter obviously unbounded unless $b^0 = 0$) will not be discussed here. Similarly, CCN are immediately solved by linear algebra techniques, and therefore are of no interest in this context. On the other end of the spectrum, in general QP is a hard problem. Actually, LIQ---linear objective function and quadratic constraints in integer variables with no finite bounds, i.e.
\[
 \textstyle
 \min \; \Big\{ \, b^0 x \;:\;
 \half x^\top Q^i x + b^i x \leq c^i \quad i \in \Mcm \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcn \, \Big\}
 \;\; ,
\]
is not only $\mathcal{NP}$-hard, but downright undecidable \cite{jeroslow}. Hence so are the ``harder'' \{C,Q\}IQ.

@@

Another relevant reformulation trick concerns the fact that, as soon as quadratic constraints are allowed, then a linear objective function can be assumed w.l.o.g.. Indeed, any Q** (C*C) problem can always be rewritten as
%
\begin{align*}
 \min \;\;
 & x^0 \\
 %
 & -\infty \leq \half x^\top Q^0 x + b^0 x \leq x^0 & \\
 %
 & c^i_l \leq \half x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\
 %
 & l_j \leq x_j \leq u_j & j \in \Mcn  \\
 %
 & x_j \in \mathbb{Z} & j \in \Mcz
\end{align*}
%
i.e., a L*Q (L*C). Hence, it is clear that quadratic constraints are, in a well-defined sense, the most general situation (cf.~also the result above about hardness of LIQ).

When a $Q^i$ is positive semidefinite (PSD), i.e., the corresponding constraint/objective function is convex, general quadratic constraints are in fact equivalent to diagonal ones. In fact, since every PSD matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, the term $\half x^\top Q^i x \equiv \half \sum_{j \in \Mcn} z^2_j$ where $z^\top = x^\top L^i$. Hence, one could think that D** problems need not be distinguished from C** ones; however, this is true only for ``complicated'' constraints, but not for ``simple'' ones, because the above reformulation technique introduces linear constraints. Indeed, while C*L (and, a fortiori, C*\{C,Q\}) can always be brought to D*L (D*\{C,Q\}), using the same technique C*B becomes D*L, which is significantly different from D*B. In practice, a diagonal convex objective function under linear constraints is found in many applications (e.g., \cite{FrFG16,FrGe06a,FrGe09a,FGGP11}), so that it still makes sense to distinguish the D*L case where the objective function is ``naturally'' separable from that where separability is artificially introduced.

### 2017-02-01 QPLIB-3.tex
update to qplib format and that pesky 1/2

In this section we present all the steps performed in order to build the
new library. In \S~\ref{subsec:instColl}, we describe the set of gathered
instances. In \S~\ref{subsec:selection} we present the features used to
classify the instances and we discuss the issues concerning the format of
the instances. Finally, in \S~\ref{subsec:final set}, we describe the
selection process used to filter the instances and we graphically present
the main features of the selected instances.

@@

To evaluate the instances we decided, for practical reasons, to use
GAMS as common platform for all the experiments involving commercial and
non-commercial solvers. For this reason, we decided to translate all
instances into the GAMS format (\texttt{.gms}). 
% In a preliminary phase, all the instances received were divided
%according to their format and subsequently translated.
%In \S~\ref{subsec:tools}
%%~\ref{subsec:CP_convex}
%the tools used
%to translate an instance from a given format to the \texttt{.gms} format
%%to translate an instance from and to a given format to the
%\texttt{.gms} format
%are described in more detail.\\
%
In addition, we have introduced a new, specific \texttt{.qplib} format. This
new format is capable of describing all the instances of the library in a sparse form.
In comparison to a more \emph{high level} format like \texttt{.gms}, the new
format presents three advantages: it is easier to read by a self-made parser, it produces smaller files, and the order of variables and constraints is fixed. See the \S~\ref{sec:format} for more details.

@@

For each instance of the starting set, we collected important characteristics
which allowed us to classify the instances into the QP categories described in
\S~\ref{sec:QPbasic}. As far as the variable types are concerned, we
collected the following information:
%
\begin{itemize}
 \item number of binary variables; % (\emph{\# bin}),
 %
 \item number of integer variables; % (\emph{\# int}),
 %
 \item number of continuous variables. % (\emph{\# cont}).
\end{itemize}
%
If at least one binary or integer variable is present, then the instance is
categorized as \emph{discrete}, otherwise it is categorized as \emph{continuous}.
As far as the objective function is concerned, we gathered the following
information:
%
\begin{itemize}
 \item percentage of positive and negative eigenvalues of the $Q^0$ matrix;
       % (\emph{\% neg eig}),
 %
 \item density of the $Q^0$ matrix (number of nonzero entries divided by the total
       number of entries). %(\emph{\% dens}).
\end{itemize}
%

@@

During the development of the library, a discussion ensued about
the expected goals that we wanted to achieve. The following four goals
where finally identified:
%
\begin{enumerate}
 \item represent as much as possible all the different categories of QP
       problems;
 %
 \item gather ``challenging'' instances, i.e., ones which can not be easily
       solved by  state-of-the-art solvers;
 %
 \item include for each of the categories a set of well-diversified
       instances;
 %
 \item obtain a set of instances which is neither too small, so as to
       preserve statistical relevance, nor too large so as to being
       computationally manageable.
\end{enumerate}
%
To achieve the aforementioned goals, we performed the following two
filters, applied in cascade.
%
\begin{itemize}
 \item \emph{First Instances Filter.}\\
       The first filter was designed to drastically reduce the number of
       instances by eliminating the ``easy'' ones. An empirical measure
       for the hardness of an instance is the CPU time needed by a
       complete solver (cf.~\S~\ref{sec:algo}) to solve it to
       global optimality. Accordingly, for each of the gathered instance we
       ran the complete solvers in GAMS, which number depends on the category
       of the instance under consideration, cf.~Table \ref{t:solvers}.
       We then filtered according to a first
       measure of computational difficulty, i.e., we discarded all instances
       that are solved by at least 30\% of the complete solvers within a time
       limit of 30 seconds.
 %
 \item \emph{Second Instances Filter.}\\
       The goal of the second filter was to eliminate ``similar'' instances.
       We carefully analysed the instances one by one, and we clustered them
       according to their features; for each cluster we kept only a few
       representatives (e.g., very similar size, same donor,\ldots). Finally, in order to only
       keep computationally challenging instances we ran a a complete solver for QGQ with a time limit of 120 seconds; all the
       instances which have been solved to proven optimality within this time limit
       were discarded.
\end{itemize}
%
In Table~\ref{tab:filters} we summarize the two filter steps, which
allowed us to identify the final set of $251$ discrete instances and
$116$ continuous instances.

@@

Speaking of nonconvexity, Figure~\ref{fig:pic_neg_eig} (left) focuses on
the instances with a quadratic objective function, ordered by the
the percentage ``hard'' eigenvalues in the $Q^0$ matrix (vertical axis), by which
we mean negative eigenvalues in case of a minimization problem and positive eigenvalues in case of a maximization problem.
The instances are mostly clustered around two values. About 25\% of the instances have a convex (if minimization) or concave (if maximization) objective function,
i.e., they have 0\% of ``hard'' eigenvalues. Among the others, a vast
majority has around 50\% of ``hard'' eigenvalues. However, instances
with high or low percentages of ``hard'' eigenvalues are present, too.

\begin{figure}\centering
  \includegraphics[width=0.49\textwidth]{pic_neg_eig.pdf}
  \includegraphics[width=0.49\textwidth]{pic_density.pdf}
  \caption{``Hard'' eigenvalues (left) and density (right) of $Q^0$ matrix for QPLIB instances with a quadratic objective function. \label{fig:pic_neg_eig}}
\end{figure}

Similarly, Figure~\ref{fig:pic_neg_eig} (right) shows the instances with a
quadratic objective function sorted according to the density of the
$Q^0$ matrix (vertical axis). The majority of the instances have either
a very low or a rather high density: indeed, about 30\% of the instances
have density smaller than 5\%, and about 30\% of the instances have
density larger than 50\%. However, also intermediate values are present.

Additional details on the instance features can be found in \S~\ref{sec:instance_details}.

>>>>

In this section we present all the steps performed in order to build the
new library. In \S\ref{subsec:instColl}, we describe the set of gathered
instances. In \S\ref{subsec:selection} we present the features used to
classify the instances and we discuss the issues concerning the format of
the instances. Finally, in \S\ref{subsec:final set}, we describe the
selection process used to filter the instances and we graphically present
the main features of the selected instances.

@@

To evaluate the instances we decided, for practical reasons, to use
GAMS as common platform for all the experiments involving commercial and
non-commercial solvers. For this reason, we decided to translate all
instances into the GAMS format (\texttt{.gms}).
% In a preliminary phase, all the instances received were divided
%according to their format and subsequently translated.
%In \S\ref{subsec:tools}
%%~\ref{subsec:CP_convex}
%the tools used
%to translate an instance from a given format to the \texttt{.gms} format
%%to translate an instance from and to a given format to the
%\texttt{.gms} format
%are described in more detail.\\
%
In addition, we have introduced a new, specific \texttt{.qplib} format. This
new format is capable of describing all the instances of the library in a sparse form.
In comparison to a more \emph{high level} format like \texttt{.gms}, the new
format presents three advantages: it is easier to read by a self-made parser, it produces smaller files, and the order of variables and constraints is fixed. See the \S\ref{sec:format} for more details.

@@

For each instance of the starting set, we collected important characteristics
which allowed us to classify the instances into the QP categories described in
\S\ref{sec:QPbasic}. As far as the variable types are concerned, we
collected the following information:
%
\begin{itemize}
 \item number of binary variables; % (\emph{\# bin}),
 %
 \item number of integer variables; % (\emph{\# int}),
 %
 \item number of continuous variables. % (\emph{\# cont}).
\end{itemize}
%
If at least one binary or integer variable is present, then the instance is
categorized as \emph{discrete}, otherwise it is categorized as \emph{continuous}.
As far as the objective function is concerned, we gathered the following
information:
%
\begin{itemize}
 \item percentage of positive and negative eigenvalues of the Hessian $Q^0$;
       % (\emph{\% neg eig}),
 %
 \item density of the Hessian $Q^0$ (number of nonzero entries divided by the total
       number of entries). %(\emph{\% dens}).
\end{itemize}
%

@@

During the development of the library, a discussion ensued about
the expected goals that we wanted to achieve. The following four goals
where finally identified:
%
\begin{enumerate}
 \item represent as much as possible all the different categories of QP
       problems;
 %
 \item gather ``challenging'' instances, i.e., ones which can not be easily
       solved by  state-of-the-art solvers;
 %
 \item include for each of the categories a set of well-diversified
       instances;
 %
 \item obtain a set of instances which is neither too small, so as to
       preserve statistical relevance, nor too large so as to being
       computationally manageable.
\end{enumerate}
%
To achieve the aforementioned goals, we performed the following two
filters, applied in cascade.
%
\begin{itemize}
 \item \emph{First Instances Filter.}\\
       The first filter was designed to drastically reduce the number of
       instances by eliminating the ``easy'' ones. An empirical measure
       for the hardness of an instance is the CPU time needed by a
       complete solver (cf.~\S\ref{sec:algo}) to solve it to
       global optimality. Accordingly, for each of the gathered instance we
       ran the complete solvers in GAMS, which number depends on the category
       of the instance under consideration, cf.~Table \ref{t:solvers}.
       We then filtered according to a first
       measure of computational difficulty, i.e., we discarded all instances
       that are solved by at least 30\% of the complete solvers within a time
       limit of 30 seconds.
 %
 \item \emph{Second Instances Filter.}\\
       The goal of the second filter was to eliminate ``similar'' instances.
       We carefully analysed the instances one by one, and we clustered them
       according to their features; for each cluster we kept only a few
       representatives (e.g., very similar size, same donor,\ldots). Finally, in order to only
       keep computationally challenging instances we ran a a complete solver for QGQ with a time limit of 120 seconds; all the
       instances which have been solved to proven optimality within this time limit
       were discarded.
\end{itemize}
%
In Table~\ref{tab:filters} we summarize the two filter steps, which
allowed us to identify the final set of $251$ discrete instances and
$116$ continuous instances.

@@

Speaking of nonconvexity, Figure~\ref{fig:pic_neg_eig} (left) focuses on
the instances with a quadratic objective function, ordered by the
the percentage ``hard'' eigenvalues in the Hessian $Q^0$ (vertical axis), by which
we mean negative eigenvalues in case of a minimization problem and positive eigenvalues in case of a maximization problem.
The instances are mostly clustered around two values. About 25\% of the instances have a convex (if minimization) or concave (if maximization) objective function,
i.e., they have 0\% of ``hard'' eigenvalues. Among the others, a vast
majority has around 50\% of ``hard'' eigenvalues. However, instances
with high or low percentages of ``hard'' eigenvalues are present, too.

\begin{figure}\centering
  \includegraphics[width=0.49\textwidth]{pic_neg_eig.pdf}
  \includegraphics[width=0.49\textwidth]{pic_density.pdf}
  \caption{``Hard'' eigenvalues (left) and density (right) of the Hessian $Q^0$ for QPLIB instances with a quadratic objective function. \label{fig:pic_neg_eig}}
\end{figure}

Similarly, Figure~\ref{fig:pic_neg_eig} (right) shows the instances with a
quadratic objective function sorted according to the density of the
Hessian $Q^0$ (vertical axis). The majority of the instances have either
a very low or a rather high density: indeed, about 30\% of the instances
have density smaller than 5\%, and about 30\% of the instances have
density larger than 50\%. However, also intermediate values are present.

Additional details on the instance features can be found in \S\ref{sec:instance_details}.
